Response to referees: Multi-target detection with application to cryo-electron microscopy

### 2019-05-29 InverseProblemsV2/reponse.tex
response to referees

\title{Multi-target detection with application to cryo-electron microscopy}

\author[a]{Tamir Bendory}
\author[b]{Nicolas Boumal} 
\author[c]{William Leeb}
\author[a,b]{Eitan Levin}
\author[a,b]{Amit Singer}

\affil[a]{The Program in Applied and Computational Mathematics, Princeton University, Princeton, NJ, USA}
\affil[b]{Department of Mathematics, Princeton University, Princeton, NJ, USA}
\affil[c]{School of Mathematics, University of
	Minnesota, Minneapolis, MN, USA }

%\date{}
\maketitle

\begin{abstract}

We consider the multi-target detection problem of recovering a set of signals that appear 
multiple times at unknown locations in a noisy measurement.
In the low noise regime, one can estimate the signals by first detecting occurrences, then clustering and averaging them.
In the high noise regime, however, neither detection nor clustering can be performed reliably, so that strategies along these lines are destined to fail.  
Notwithstanding, using autocorrelation analysis, we show that the impossibility to detect and cluster signal occurrences in the presence of high noise does not necessarily preclude signal estimation.
Specifically, to estimate the signals, we derive simple relations between the autocorrelations of the observation and those of the signals. These autocorrelations can be estimated accurately at any noise level given a sufficiently long measurement. 
%To recover the signals from the observed autocorrelations, we (heuristically) solve a set of polynomial equations through nonlinear least-squares. 
To recover the signals from the observed autocorrelations, we  solve a set of polynomial equations through nonlinear least-squares. 
We provide analysis regarding well-posedness of the task, and demonstrate numerically the effectiveness of the method in a variety of settings.

The main goal of this work is to provide theoretical and numerical support for a recently proposed framework to image 3-D structures of biological macromolecules using cryo-electron microscopy in extreme noise levels. % ~\cite{bendory2018toward}. -- not references in the abstract, except if cited in full; but we don't want to write "arxiv" here.

\end{abstract}

\section{Introduction} \label{sec:intro}

We consider the \emph{multi-target detection} problem of recovering a set of $K$ signals that appear 
multiple times at unknown locations in a noisy measurement.
Let ${x_1,\ldots,x_K\in\RL}$ be the sought signals and let $y\in\RN$ be the observed data, where we assume $N$ is  much larger than $L$. 
Let  $s[i]$ count the number of signal occurrences whose first entry is positioned at $y[i]$. Each of those $s[i]$ signals is chosen among $x_1, \ldots, x_K$ according to some (possibly unknown) distribution over $\{1,\ldots,K\}$. 
If signal occurrences overlap, they interfere additively. %We think of those events as being rare. 
With additive white Gaussian noise, the measurement model can be written as 
\begin{align} 
	y  &=  \sum_{k=1}^K s_k \ast x_k + \varepsilon, &\varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
	\label{eq:model}
\end{align}
where $\ast$ denotes linear convolution, and $s_k[i]$ indicates the number of occurrences of $x_k$ starting at $y[i]$, so that $s =  s_1+\cdots+s_K$. Explicitly, with zero-based indexing, % and zero-padding out of bounds,
\begin{align*}
	y[i] & = \sum_{k=1}^{K} \sum_{j = 0}^{L-1} s_k[i-j] x_k[j] + \varepsilon[i].
\end{align*}
%We discuss generative models for $s_1, \ldots, s_K$ below.
%
The goal is  to estimate $x_1,\ldots,x_K$ from $y$. % while $s$ is the \emph{nuisance variable} of the problem.
%We name this problem \emph{multi-target detection}.
In parts of the paper, we focus on the case $K = 1$, called the \emph{homogeneous} case; the case $K \geq 2$ is called \emph{heterogeneous}.
This idealized setup appears in several scientific applications, including structural biology~\cite{bendory2018toward} (as we detail below), spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive}, and system identification~\cite{ljung1998system}.

In the low noise regime (small $\sigma$), a valid strategy is to first detect the signal occurrences in $y$ (that is, estimate $s$), cluster them (that is, separate $s$ into $s_1,\ldots,s_K$), and solve  standard deconvolution problems.
Crucially, we focus on the high noise regime, where \emph{reliable detection of signal occurrences is  impossible}~\cite{bendory2018toward,aguerrebere2016fundamental}.
This limitation does not, however, preclude estimation of the signals $x_1,\ldots,x_K$, as we show in this paper. In this setting, we refer to  $s_1,\ldots,s_K$  as \emph{nuisance variables}: knowing them would certainly help, but we do not aim to estimate them.

In order to recover the signals in the high noise regime, we use autocorrelation analysis.
At any noise level, the autocorrelations of the observation can be estimated to any desired accuracy for sufficiently large  $N$. 
This computation is straightforward and requires only one pass over the data.
The underlying principle is to relate the autocorrelations of the observation $y$ to the autocorrelations of $x_1,\ldots,x_K$.

Below we describe two generative models for $s$.  
In these models, the relationship between the autocorrelations of $y$ and those of $x_1,\ldots,x_K$ depends on $s_1,\ldots,s_K$ only through their expected sums, that is, the expected total number of occurrences of each signal.
To estimate the signals and occurrence counts from the computed autocorrelations, we solve a nonlinear least-squares problem as explained in Section~\ref{sec:numerics}.

The multi-target detection problem  is an instance of  
\emph{blind deconvolution}---a longstanding problem arising in a variety of engineering and scientific applications, such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, to name a few. 
Different variants of the blind deconvolution problem have been thoroughly analyzed recently~\cite{ahmed2014blind,li2016identifiability,li2016rapid,lee2017blind,ling2017blind,kuo2019geometry}. In clear contrast to multi-target detection, these works focus on the low noise regime and aim to estimate both unknown signals (in our setting, this means estimating both $x_k$'s and $s_k$'s).

\subsection*{Models for target distribution}

We consider two  models for the distribution of signal occurrences in the observation, that is, for $s_1, \ldots, s_K$.

\paragraph{The well-separated model.}

As a first setup, we allow any generative model for $s$ which meets the following separation requirement: $s$ is binary, and
\begin{equation}
\textrm{If } s[i] = 1 \textrm{ and } s[j] = 1 \textrm{ for } i \neq j, \textrm{ then } |i - j| \geq 2L-1.
\label{eq:spacing}
\end{equation}
In words: the starting positions of any two occurrences  must be separated by at least $2L-1$ positions, so that their end points are necessarily separated by at least $L-1$ signal-free (but still noisy) entries in the data.
Furthermore, we require that the last signal occurrence in $y$ is also followed by at least $L-1$ signal-free entries.
This property ensures that correlating $y$ with versions of itself shifted by at most $L-1$ entries does not involve correlating distinct signal occurrences. Once $s$ is determined, for each position $i$ such that $s[i] = 1$, one of the signals $x_k$ is selected independently at random, and accordingly we set $s_k[i] = 1$. As a result, the only properties of $s_1, \ldots, s_K$ that affect the autocorrelations of $y$ (for shifts up to $L-1$) are the total number of occurrences of the distinct signals: their individual and relative locations do not intervene. We detail this in Section~\ref{sec:AC_analysis}.

%\revised{A follow-up paper~\cite{lan2019multi} extends this model by allowing an arbitrary spacing between signal occurrences, while excluding overlapping occurrences}.

\paragraph{The Poisson model.}

If the separation condition is violated, more knowledge about the location distribution is necessary to disentangle the autocorrelations of $y$. To that effect, we analyze a Poisson generative model.

Specifically, for each position $i$, the number $s[i]$ of signal occurrences starting at that position is drawn independently from a Poisson distribution with parameter $\gamma/L$, that is, ${s[i] \overset{i.i.d.}{\sim} \Poisson(\gamma / L)}$ for some parameter $\gamma > 0$. Then, $s[i]$ is split into $s_1[i] + \cdots + s_K[i]$ by selecting $s[i]$ signals among $x_1, \ldots, x_K$, independently at random  following a fixed distribution over $\{1,\ldots,K\}$.
It is possible for more than one occurrence of the same $x_k$ to start at position~$i$.
As for the well-separated model, the autocorrelations of $y$ under this model depend weakly on $s$ and $s_1, \ldots, s_K$, essentially through the parameter $\gamma$: see Section~\ref{sec:AC_analysis}.

\subsection*{Extensions} \label{sec:extensions}

Extending the problem setup and autocorrelation analysis to signals in more than one dimension is straightforward: see the discussion in Section~\ref{sec:high_dimensions} and numerical experiments in Section~\ref{sec:numerics}.

Likewise, it is easy to extend the model to situations where the signal occurrences are sampled from a general distribution rather than from a finite set of choices $x_1, \ldots, x_K$. The finite setup corresponds to the following distribution for signal occurrences:
\begin{align}
	x \sim \sum_{k=1}^K \pi_k \delta(x - x_k),
	\label{eq:heterogeneousdistribution}
\end{align}
where $\delta(x - x_k)$ is a Dirac delta (a point mass) located at $x_k$, and $(\pi_1, \ldots, \pi_K)$ encodes the discrete distribution over $\{1, \ldots, K\}$.
In the generalized setup, the goal is to estimate the distribution (possibly defined by a finite set of parameters). In particular, this allows for continuous distributions of targets. We adopt this perspective when deriving the autocorrelations in Section~\ref{sec:AC_analysis}.

In the next section, we show how this flexibility allows us to model an important imaging problem in structural biology.

\section{Connection with single-particle reconstruction via cryo-electron microscopy} \label{sec:cryoem}

Cryo-electron microscopy (cryo-EM)  has recently joined X-ray crystallography and nuclear magnetic resonance (NMR) spectroscopy as a high-resolution structural method for biological macromolecules~\cite{frank2006three,kuhlbrandt2014resolution,bartesaghi20152}.
%In contrast to X-ray and NMR which aggregate information from an ensembles of
%particles, single particle cryo-EM produces images of individual particles and thus can, in principle, elucidate multiple  structures simultaneously.
%In addition, it does not require the formation of crystalline arrays of macromolecules.
In a cryo--EM experiment, biological samples (e.g., macromolecules, viruses) are rapidly frozen in a thin layer of vitreous ice.
The microscope produces 2-D tomographic images of the samples embedded in the ice, called \emph{micrographs}. Each micrograph contains multiple tomographic projections of the samples at unknown locations and under unknown viewing directions. 
Importantly, the electron dose must be kept low to mitigate  radiation damage, inducing high noise levels.
The goal is to reconstruct 3-D models of the molecular structures from the micrographs. 
Since  cryo-EM produces images of individual particles, it can elucidate multiple  structures simultaneously.
This is in clear contrast to  X-ray and NMR, which aggregate information from an ensembles of
particles.

Considering the extensions described in the previous section, we can phrase a simplified generative model for micrographs (the observation $y$) within our framework. Specifically, locations are chosen in the 2-D plane of the image, corresponding to $s$: this is where the molecules are fixed in the plane of the ice layer. At each selected location, a 2-D tomographic projection of a molecule (a signal occurrence) is added in the observation $y$. This signal is drawn from a probability distribution described by a discrete number of parameters which correspond to the sought 3-D structure, as follows.

The 3-D structure $V$ (the target parameter) can be expanded into a linear combination of basis functions (for example, spherical harmonics for the spherical part and Bessel functions for the radial part): the coefficients of this expansion are the unknowns. Then, a rotation $R_\omega$ is applied to the volume (to model viewing directions) according to a (possibly unknown) distribution of $\omega$ over $SO(3)$ (the group of 3-D rotations). With tomographic projection denoted by $P$, $x$ is a random signal, related to the distribution of $\omega$ through: 
%we have the following probability distribution for particles in the micrograph:
\begin{align}
	x & = P(R_\omega V).
	\label{eq:cryoEM}
\end{align}
By further allowing $V$ itself to be a random signal as well, this model can also encode a mixture of structures in the biological sample. This mixture might also be continuous, corresponding to continuous conformational variability.

All contemporary methods for single particle reconstruction using cryo-EM  split the reconstruction procedure into two main  stages.
The first stage, called \emph{particle picking},  detects and extracts the particle projections from the micrographs. Given the projections, the second stage aims to reconstruct a 3-D model of the molecular structure, usually using an expectation-maximization algorithm~\cite{scheres2012relion}. 
%The quality of the reconstruction eventually hinges on the quality of the particle picking stage.
Crucially, reliable detection of individual particles is impossible in a highly noisy environment. This fact has been recognized early on by the cryo-EM community. 
Particularly, in~\cite{henderson1995limitations,glaeser1999electron}, it was reasoned that particle picking is impossible for molecules below a certain weight (below~$\sim$50 kDa). 
%The unique issues raised by small particles have been mitigated by recent technological advances in the field, including the use of Volta phase plates~\cite{khoshouei2017cryo,liang2017phase} and scaffolding cages~\cite{liu2018nearatomic}.
%Despite this progress, detecting small molecules in the micrographs remains a challenge.

Even if particle picking is feasible, procedures may be affected by \emph{model bias}.
Particle picking algorithms are typically based on correlating the micrograph with several templates. Areas highly correlated with some of the templates are assumed to contain projections: these yield the picked particles. 
Clearly, the picked particles depend heavily on the chosen templates: different templates may lead to different picked particles, hence, which is more problematic, to different 3-D structure reconstructions.
The cryo-EM community is well aware of this potential pitfall~\cite{vanheel1992correlation,henderson2013avoiding,vanheel2013finding}, which was notably exemplified \revised{by} the ``Einstein from noise'' experiment~\cite{shatsky2009method}.

A recent work of the authors suggests a methodology to bypass particle picking and  reconstruct the 3-D structure directly from the micrograph~\cite{bendory2018toward}.
Based on autocorrelation analysis, it was shown that---at least in principle---the limits high noise regimes 
impose on particle picking  do not necessarily translate into limits on 3-D reconstruction.
The main goal of  the present paper is to provide theoretical and numerical support for this approach, in a simplified setting.

In the next section, we introduce autocorrelation analysis in detail, focusing on the multi-target detection model. 
We mention that similar ideas in  cryo-EM can be traced back to a seminal paper of Zvi Kam~\cite{kam1980reconstruction}. Kam proposed autocorrelation analysis for \mbox{3-D} reconstruction, under the assumption of picked, perfectly centered, particles. Kam's method has been extended and used in X-ray free electron lasers (XFEL) and cryo-EM~\cite{liu2013three,kurta2017correlations,levin20173d,von2018structure}.  
In order to investigate the computational and statistical properties of Kam's method, a series of papers have studied a simplified model, 
called   \emph{multi-reference alignment}~\cite{bandeira2014multireference,bendory2017bispectrum,bandeira2017optimal,perry2017sample,bandeira2017estimation,abbe2017multireference}.
We follow the same line of research by considering the multi-target detection as an abstraction to the application of reconstructing 3-D structures directly from the micrograph~\cite{bendory2018toward}.

\section{Autocorrelation analysis} \label{sec:AC_analysis}

In what follows, we consider autocorrelations of both the observation $y$ and of the signal occurrences in $y$. As per our discussion of extensions, the signal occurrences may be sampled from a discrete set $\{x_1, \ldots, x_K\}$ (as in~\eqref{eq:heterogeneousdistribution}), or from a more general distribution. Accordingly, we define autocorrelations broadly for a random signal $z$ of length $M$. For our purposes, this will be applied both to signal occurrences (of length $L$) and to $y$ (of length $N$).

For a random signal $z\in \R^M$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \E_z\left\{\frac{1}{M} \sum_{i=-\infty}^{\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\},
\label{eq:ac_general}
\end{equation}
where the expectation is taken with respect to the distribution of $z$. Indexing out of bounds is zero-padded, that is, $z[i] = 0$ for $i$ out of the range $0, \ldots, M-1$.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align}  
a_z^1 & = \E_z\left\{\frac{1}{M} \sum_{i=0}^{M-1} 
z[i]\right\}, \nonumber\\
a_z^2[\ell] & = \E_z\left\{\frac{1}{M} \sum_{i = \max\{0, -\ell\}}^{M-1 + \min\{0, -\ell\}} z[i]z[i+\ell]\right\}, \label{eq:ac_special} \\
a_z^3[\ell_1,\ell_2] & = \E_z\left\{\frac{1}{M} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{M-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]\right\}. \nonumber
\end{align}
Since  autocorrelations depend only on the differences between indices, they obey the following symmetries: $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$
%In what follows, $x$ is the random variable corresponding to signal occurrences in $y$.
% In particular, for the model~\eqref{eq:model}, 
%$x=x_k$ with probability $\pi_k$.
In particular, for
$x$ sampled from $\{ x_1, \ldots, x_K \}$ with probabilities $(\pi_1, \ldots, \pi_K)$ as in~\eqref{eq:heterogeneousdistribution}, the autocorrelations of $x$ are given in explicit form in terms of those of the deterministic signals $x_1, \ldots, x_K$ as:
\begin{align}
	a_x^q  & = \sum_{k=1}^{K} \pi_ka_{x_k}^q.
	\label{eq:mixedautocorr}
\end{align}
Explicit expressions of the autocorrelations for the more involved model of  cryo-EM~\eqref{eq:cryoEM} are given in~\cite{bendory2018toward}.

We are given one observation (one realization) of $y$. Thus, we cannot compute the autocorrelations of $y$ exactly as they involve taking an expectation against the distribution of $y$. However, by the law of large numbers, as $N$ grows to infinity  the empirical autocorrelations of $y$ almost surely (a.s.) converge to the actual (population) autocorrelations of $y\in \RN$, that is, % (ergodicity)
\begin{align}
\lim_{N \to \infty} \frac{1}{N} \sum_{i=-\infty}^{\infty} y[i]y[i+\ell_1]\cdots y[i+\ell_{q-1}] \aseq a_y^q[\ell_1, \ldots, \ell_{q-1}].
\end{align}
This provides a concrete means of estimating the quantities $a_y^ q$. In the remainder of this section, we relate the observables $a_y^q$  to the unknowns $a_x^q$, first under the well-separated model, then under the Poisson model.

\subsection{Autocorrelations under the well-separated model}

Under the separation condition~\eqref{eq:spacing}, the relation between autocorrelations of the observation $y$ and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $\vert s\vert = \sum_i s[i]$ denote the number of signal occurrences in $y$, and let
\begin{equation}
\gamma  = \frac{|s| L}{N}.
\label{eq:gamma}
\end{equation}
This $\gamma$ is the fraction of entries of $y$ occupied by signal occurrences. %Henceforth, we call it the signal density.
The separation condition  imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$. 
For the heterogeneous model~\eqref{eq:model},  one can define
\begin{equation}
\gamma_k  = \frac{|s_k| L}{N},
\label{eq:gamma_k}
\end{equation}
where $\vert s_k\vert = \sum_i s_k[i]$ so that $\gamma=\sum_k \gamma_k$.

Owing to the separation condition, when correlating $y$ with shifted versions of itself for shifts in $0, \ldots, L-1$, any given occurrence of $x$ in $y$ is only ever correlated with itself, and never with another occurrence. As a result, the autocorrelations of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the density $\gamma$ (which is a weak dependence on the support signal $s$). Specifically, we show the following identities in Appendix~\ref{sec:autocorrelation_computation}:
\begin{align} 
	a_y^1 & = \gamma a_{x}^1, \label{eq:mean_micrograph} \\
	a_y^2[\ell] & = \gamma a_{x}^2[\ell] + \sigma^2\delta[\ell], \label{eq:ac2_micrograph}\\
	a_y^3[\ell_1,\ell_2] & = \gamma a_{x}^3[\ell_1,\ell_2]  + \sigma^2\gamma a_{x}^1  \big(\delta[\ell_1]+\delta[\ell_2]
	+\delta[\ell_1-\ell_2]\big), \label{eq:ac3_micrograph}
\end{align}
where $\delta[0] = 1$ and $\delta[\ell \neq 0] = 0$, and indices $\ell, \ell_1, \ell_2$ are in the range $0 \leq \ell \leq L-1$. Terms proportional to $\sigma^2$ are due to noise. If $\sigma$ is known, they can be handled easily. If $\sigma$ is unknown, one can either estimate it form the data, or one can ignore the few entries of the autocorrelations that are affected by $\sigma$---one in $a_y^2$ and $3L-2$ in $a_y^3$, a relatively small number in both cases.

We show in Section~\ref{sec:theory_homogeneous_well_separated} 
that $x,\gamma$ and $\sigma$ can be identified uniquely from the observed autocorrelations for the homogeneous case $K = 1$ (all signal occurrences are the same).

\subsection{Autocorrelations under the Poisson model} \label{sec:ac_poisson}

In this section, we give the expressions for the autocorrelations of the observed signal $y$ under the Poisson model. %In this model, each location $k$, $1 \le k \le N-L+1$ is the starting point of a Poisson-distributed number of random signals. 
%More precisely, letting $s[k]$ be the number of signals starting at location $k$, $s[k] \overset{iid}{\sim} \Poisson(\gamma / L)$.
 We first note that the expected total number of signals occurring in $y$ is equal to $(N - L + 1)\gamma / L$, and consequently the total lengths of all signals (including overlaps) divided by $N$ is equal to
%
\begin{align}
\frac{N - L + 1}{N} \cdot \frac{\gamma  }{L}  \cdot L 
\quad \overset{N\to\infty}{\longrightarrow} \quad \gamma.
\end{align}
%
Therefore, in the large $N$ limit, similarly to the role of $\gamma$ in the well-separated model~\eqref{eq:gamma}, the parameter $\gamma$ may be interpreted as the signal density.

In Appendix~\ref{sec:proof_prop_poisson}, we prove the following expressions for the autocorrelations of $y$ under the Poisson model. As in the well-separated model, the observed autocorrelations do not depend on individual occurrences of the signal, but only on the distribution of $x$ itself, and the parameters $\gamma$ and $\sigma$: 
%
\begin{align}
	a_y^1  & =  \gamma a_{x}^1, \label{eq:mean_micrograph2} \\
	a_y^2[\ell] & = \gamma a_x^2[\ell] + \sigma^2 \delta[\ell] + (\gamma a_x^1)^2, \label{eq:ac2_micrograph2} \\
	a_y^3[\ell_1,\ell_2]  & =  \gamma a_x^3[\ell_1,\ell_2] + \sigma^2 \gamma a_x^1 \big(\delta[\ell_1]+\delta[\ell_2] +\delta[\ell_1-\ell_2] \big) \nonumber \\
						  & \qquad + (\gamma a_x^1)^3 + \gamma a_x^1 \cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2] + \gamma a_x^2[\ell_2-\ell_1]). \label{eq:ac3_micrograph2}
\end{align}
Note that from those autocorrelations, it is easy to retrieve the autocorrelations of the well-separated model for all entries unaffected by $\sigma$. If $\sigma$ is known, then it is true for all entries.

In the homogeneous case ($K=1$), we show in Section~\ref{sec:theory_homogeneous_poisson} that $a_y^1,a_y^2$ and $a_y^3$ identify uniquely the signal $x$, the Poisson parameter $\gamma$ and the noise level $\sigma$ for generic $x$.

\section{Theory}

We begin this section by showing that, in the homogeneous case ($K=1$), under both the well-separated and the Poisson models, the first three observed autocorrelations  identify the (deterministic) signal $x$, the density parameter $\gamma$, and the noise level $\sigma^2$.  Then, for the heterogeneous case ($K \geq 2$), we bound from above the number $K$ of signals that can be recovered from those autocorrelations as a function of the signal length $L$. Finally, we briefly discuss multi-dimensional signals.

Before that, we start by showing that in the homogeneous case a deterministic signal $z\in\RL$ is identified uniquely by its second and the third autocorrelations.
 Indeed, assuming $z[0]$ and $z[L-1]$ are nonzero, we can recover $z$ explicitly by 
\begin{equation}
%
z[k]  = \frac{z[0]z[k]z[L-1]}{z[0]z[L-1]} = \frac{a_z^3[k,L-1]}{a_z^2[L-1]},
\label{eq-uniqueness}
%
\end{equation}
for $k = 0, \ldots, L-1$. If $z[0]$ or $z[L-1]$ are equal to zero, then $a_z^2[L-1] = 0$ and we can use that indication to shrink $L$.
This proves the following useful fact:
\begin{proposition} \label{prop:uniqueness}
	%
	A deterministic signal $z\in\RL$ is determined uniquely by  $a_z^2$ and $a_z^3$. 
\end{proposition}

Note that the procedure described in~\eqref{eq-uniqueness} is not numerically stable: if $z[0]$ or $z[L-1]$ are close to 0, recovery of $z$ is sensitive to errors in the autocorrelations. In practice, we recover $z$ by fitting it to its autocorrelations using a nonconvex least-squares procedure, which is empirically robust to additive noise. In prior work, we have observed similar phenomena for the related problem of multi-reference alignment~\cite{bendory2017bispectrum,boumal2017heterogeneous,abbe2017multireference}.

\subsection{Guarantees for the homogeneous well-separated model} \label{sec:theory_homogeneous_well_separated}

The observed moments $a_y^1,a_y^2$ and $a_y^3$ under the well-separated model do not immediately yield the autocorrelations  of the signal $x$; rather, the two are related by the noise level $\sigma$ and the signal density $\gamma$. We will show, however, that all parameters---$x, \gamma$ and $\sigma^2$---are still identified uniquely by the observed moments of $y$.

First, we observe that if  the noise level $\sigma$ is known, generally, one can estimate $\gamma$ from the first two moments of the micrograph.
The other direction is true as well: if $\gamma>0$ is known, then one can estimate $\sigma^2$.
 The proof is provided in Appendix~\ref{sec:proof_prop_gamma}.

\begin{proposition} \label{prop:gamma}
	Assume the separation condition~\eqref{eq:spacing} holds and $K=1$ (all signal occurrences in $y$ are identical, up to noise). If the mean of $x$ is nonzero, then  
	\begin{equation}
	\gamma  = \frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2},
	\end{equation}
	meaning $\gamma$ can be determined from $\sigma$ (and vice versa) using the observables $a_y^1, a_y^2$.
\end{proposition}

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ can be determined simultaneously. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
has Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements. The proof is provided in Appendix~\ref{sec:proof_prop_gamma_sigma}.

\begin{proposition} \label{prop:gamma_sigma}
	Assume $L \geq 3$, $K=1$ and assume that the separation condition~\eqref{eq:spacing} holds. 
	Then, the observed autocorrelations $a_y^1,a_y^2$ and  $a_y^3$ determine the ratio $\gamma$ and noise level $\sigma$ uniquely for a generic signal $x$. If $\gamma > \frac{1}{4}$, then this holds for any signal $x$ with nonzero mean. 
\end{proposition}

From Propositions~\ref{prop:uniqueness} and~\ref{prop:gamma_sigma} we   deduce the following:
\begin{corollary}
	Assume $L \geq 3$ and $K=1$.
	Under the separation condition~\eqref{eq:spacing}, the signal $x$, the ratio $\gamma$, and the noise level $\sigma$ are determined from the first three autocorrelation functions of $y$ if either the signal $x$ is generic, or $x$ has nonzero mean  and $\gamma > \frac{1}{4}$.
\end{corollary}
%
%As a side note, under the separation condition, the length $L$ of the signal can also be determined from the autocorrelations in the asymptotic regime, by inspection of the support of $a_y^2$.

\subsection{Guarantees for the homogeneous Poisson model} \label{sec:theory_homogeneous_poisson}

Similarly to the homogeneous well-separated model, the observed autocorrelations under the Poisson model identify $x,\gamma$ and $\sigma^2$ uniquely.
Opposed to Proposition~\ref{prop:gamma_sigma}, these quantities can be computed explicitly. 
In Appendix~\ref{sec:proof_uniqueness_poisson} we prove the following result:

\begin{proposition} \label{prop:uniqueness_poisson}
Under the homogeneous Poisson model, the signal $x$, the noise level $\sigma^2$ and the Poisson parameter $\gamma$ are identified uniquely \revised{from the observed autocorrelations}. 
In particular,
	\begin{align}
\gamma &= L \frac{(\gamma a_y^1) (a_y^2[1] - (a_y^1)^2)}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]}, \\
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2,
\end{align}
where 
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
&= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}

\end{proposition}

\subsection{Elementary limitations of the heterogeneous case}
\label{sec:heterogeneity}

In the heterogeneous model~\eqref{eq:model}, the unknowns are $K$ signals of length $L$, together with their densities $\gamma_1, \ldots, \gamma_K$~\eqref{eq:gamma_k} (equivalently: the distribution $\pi$ and overall density $\gamma$) and possibly the noise level $\sigma$. To estimate these parameters, we must collect at least as many independent equations. Within our framework, polynomial equations are provided by the observable autocorrelations, which correspond to mixed autocorrelations of the unknowns as per~\eqref{eq:mixedautocorr}. In this section, following~\cite{boumal2017heterogeneous}, we count how many equations the first three autocorrelations may provide in the best case (discounting symmetries). This leads to a straightforward information-theoretic upper bound on the number $K$ of signals which can be estimated, as a function of $L$. This is only an upper bound, though a bound of the same type was shown to be tight in a similar setting~\cite{bandeira2017estimation}.
The counting is based on the autocorrelations of the well-separated model. For entries independent of $\sigma$, the autocorrelations of the Poisson process contain the same information as those of the well-separated model. If $\sigma$ is known, this holds true for all entries; see Section~\ref{sec:ac_poisson}.

The first-order autocorrelation $a_y^1$~\eqref{eq:mean_micrograph} provides one equation. For second-order autocorrelations $a_y^2[\ell]$~\eqref{eq:ac2_micrograph}, if $\sigma$ is known we obtain $L$ equations with $\ell$ ranging from 0 to $L-1$. If $\sigma$ is unknown, we may disregard $a_y^2[0]$ (the only entry affected by $\sigma$) and still collect $L-1$ equations. Similarly, for third-order autocorrelations, $a_y^3[\ell_1, \ell_2]$~\eqref{eq:ac3_micrograph} with $0 \leq \ell_1, \ell_2 \leq L-1$ such that $\ell_2 \leq \ell_1$ includes all relevant entries for our purpose (this accounts for symmetries), providing $\frac{(L+1)(L+2)}{2}-2$ equations in total. 
If we further exclude any entries such that $\ell_1, \ell_2$ or $\ell_1 - \ell_2$ are zero to avoid the need to estimate $\sigma$, there are $\frac{(L-1)(L-2)}{2}$ remaining entries.

Hence, if $\sigma$ is known we collect
\begin{align*}
	1 + L + \frac{(L+1)(L+2)}{2}-2 = \frac{1}{2} L (L+5)
\end{align*}
equations, while if it is unknown and we choose not to estimate it, then we collect
\begin{align*}
	1 + (L-1) + \frac{(L-1)(L-2)}{2} = \frac{1}{2} L (L-1) + 1
\end{align*}
equations in total. Of course, there may be redundancy in these equations: we aim only to provide an upper bound. %under the assumption that these equations are independent.

>>>>

\title{Response to referees: Multi-target detection with application to cryo-electron microscopy}

\author{Tamir Bendory, Nicolas Boumal, Will Leeb, Eitan Levin, and Amit Singer}

\date{}
\maketitle

We thank the editor and the referees for reading our paper. Below, we provide a detailed, one-to-one response to all comments.

\section{Referee 1}

\noindent \textbf{Comment:} The only suggestion I have is to include a brief discussion of validation - the experiments here are from simulated data with known signals. If this technology were ever to be adopted in experimental practice, some form of statistical validation will be essential.\\

\noindent \textbf{Response:} The question of statistical validation is of great importance in any scientific field, including single particle reconstruction using cryo-EM. %This question is more general that the specific method applied.
Unfortunately,  there is  no rigorous method to validate a structure that was constituted using cryo-EM. In practice, researchers in the field use several useful heuristics, such as  comparing the elucidated structure with structures that were reconstructed by other imaging modalities (e.g., X-ray crystallography), or different statistical techniques (e.g., expectation-maximization). The validation problem is even more severe in the low SNR regime (the main interest of this paper) in which visual assessment of the data might be impossible. Our method does not offer insights for this question.

While validation is beyond the scope of this paper, autocorrelation analysis provides a flexible framework to describe reach generative models of the data. We added a paragraph in the summary to explain this issue.

\section{Referee 3}

\noindent \textbf{Comment:} An important point in practical image reconstruction, especially in modalities like phase retrieval, is that precise knowledge of the support of the image makes the reconstruction problem considerably easier.  Beyond that, having an image with a sharply delineated, known boundary is, in fact, a huge advantage. Such images are used in the examples in this paper, along with an exact knowledge of their support.  In most realistic imaging applications such information is not available, and this renders these problems inherently more difficult, regardless of the algorithm used. Some mention of this fact should appear in this paper.\\

\noindent \textbf{Response:} Section 4.4 discusses the application of our approach to 2-D images. In this section, we underscore the importance of the knowledge of the exact support for phase retrieval as was shown in [9]. In our setup, if the support cannot be estimated accurately, the solution is to consider the third-order autocorrelation. That being said, as we mention in Section 4, the support of the signal (at least in 1-D) can be determined from the second-order autocorrelation if enough data is provided. \\

\noindent \textbf{Comment:} Though it is clarified in the proof of Prop. 4.5 it is not immediately clear that  the right hand sides of formulae (4.3), (4.4), (4.5) are observable quantities. A remark, after the statement of the proposition, clarifying this  point would be useful.\\

\noindent \textbf{Response:} We added a remark to the proposition to make this point clear.\\

\noindent \textbf{Comment:}  I found it difficult to understand what Figure 2 is demonstrating. Perhaps it is a result of the poor rendering of the grey-scale.
\\

\noindent \textbf{Response:} We find this figure important to inspect (empirically) how many signals can be estimated simultaneously from a mix of their autocorrelations. We agree that this figure should be visualized in a colored version of the manuscript. We used a similar figure for the same purpose (but for a different problem) in a previous publication: Boumal, Nicolas, et al. ``Heterogeneous multireference alignment: A single pass approach.'' \emph{2018 52nd Annual Conference on Information Sciences and Systems (CISS)}. IEEE, 2018.
